Assessment and Optimization of Nowcast Measures for Trend Detection

### 2024-02-12 sections/02-notation.tex
Add labels to equations; add paragraph breaks

The trending detection of time series, as motivated in the introduction, is fundamentally interesting for the evaluation and comparison of methods from three areas: measurement technology, forecasting and nowcasting. This paper relates to numerical data.\\\\
Measurement aims to obtain accurate and reliable data about the current state of a system. Often the question is how to compare one method of measurement of a variable with another. For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard) for the same variable in the same period of time. The measured values for a measurement method are then available in a time series form, e.g. can be noted with $(x_t)^T_{t=0}$. To detect the systematic difference between two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero\cite{watson2010method}. In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (cf. \cite{lawrence1989concordance,koo2016guideline,}). Furthermore, graphical tools were developed, such as the
Bland-Altman diagram, which visualises the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}). \\\\
Forecasting is focused on predicting the future based on historical data and its pattern. We note in the following the real realisations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable  can be organised in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$. In practice, a forecasting method is typically evaluated by a forecast accuracy, which is defined by a certain loss function and evaluates the difference between the forecast and the actual outcome of the target variable. A comparison of several forecasters can then be made, for example, by ranking the forecast accuracies of the forecasters. In \cite{hyndman2006another} an overview of various forecasting accuracies used in practice, such as scale-dependent measures like Root Mean Square Error (RMSE) or measures based on percentage errors like Root Mean Square Percentage Error (RMSPE), is presented and discussed. \\\\
Nowcasting (also short-Term-Forecasting), on the other hand, has its origins in meteorology and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (cf. \cite{browning1989nowcasting,schmid2019nowcasting}). Nowcasting focuses on predictions for the present, the immediate future and the recent past, and is now being used in other fields, such as the economy, where important statistics on the current economic situation are available only after a considerable time lag (cf. \cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioural factors (cf. \cite{wu2021nowcasting}). Applications of nowcasting methods for econometrics can be found here in \cite{giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic} or for predictions of epidemic development here in \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}. When predicting values in the present, it is usual that the true values of the recent past are not available, so that only an estimate of these values is provided by the nowcast. Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits for $\tau$ less than or equal $t$. Similar to forecasting, nowcasters are often evaluated and  compared in the literature on the basis of performance measures such as the root mean absolute error (cf. \cite{gunther2021nowcasting}).\\\\
All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods takes into account the local absolute differences in different ways, but does not directly consider the influence of the sign of the differences. To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or prediction with continuous outcomes. Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see chapter \ref{cap: trending}). In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting or nowcasting methods are considered.\\\\
For measurement:
\[\diffxl = (x_{t+l} -x_{t})^T_{t=0}  \]
For forecasting: 
\[\diffxl = (x_{t,t+l} -y_{t})^T_{t=0}  \]
For nowcasting:
\[
\diffxl = 
\begin{cases} 
(x_{t,t+l} -x_{t,t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
(x_{t,t+l} -y_{t})^T_{t=0}  & \text{else}.
\end{cases}
    \]

>>>>

The trending detection of time series, as motivated in the introduction, is fundamentally interesting for the evaluation and comparison of methods from three areas: measurement technology, forecasting and nowcasting. This paper relates to numerical data.

Measurement aims to obtain accurate and reliable data about the current state of a system. Often the question is how to compare one method of measurement of a variable with another. For this purpose, measured values can be recorded with the two measurement methods to be compared (one can be the gold standard) for the same variable in the same period of time. The measured values for a measurement method are then available in a time series form, e.g. can be noted with $(x_t)^T_{t=0}$. To detect the systematic difference between two methods, a paired t-test can be performed on the pairwise difference of the two series of measurements to test the null hypothesis that the true mean difference is zero\cite{watson2010method}. In addition, various indices such as the interclass correlation coefficient or Lin's concordance correlation, which were originally introduced as a modification of the Pearson correlation, have been proposed in practice to check the reproducibility of the measurement or to compare different measurement methods (cf. \cite{lawrence1989concordance,koo2016guideline,}). Furthermore, graphical tools were developed, such as the
Bland-Altman diagram, which visualises the differences between the methods in relation to their mean value (cf. \cite{bland1986statistical}).

Forecasting is focused on predicting the future based on historical data and its pattern. We note in the following the real realisations of the target value with $(\mathbf{y})^T_{t=0}$ and the prediction data for a target variable  can be organised in such a way that $x_{t,\tau}$ the prediction of one predictor for time $\tau$ at time $t$ provided that $\tau$ is greater than $t$. In practice, a forecasting method is typically evaluated by a forecast accuracy, which is defined by a certain loss function and evaluates the difference between the forecast and the actual outcome of the target variable. A comparison of several forecasters can then be made, for example, by ranking the forecast accuracies of the forecasters. In \cite{hyndman2006another} an overview of various forecasting accuracies used in practice, such as scale-dependent measures like Root Mean Square Error (RMSE) or measures based on percentage errors like Root Mean Square Percentage Error (RMSPE), is presented and discussed.

Nowcasting (also short-Term-Forecasting), on the other hand, has its origins in meteorology and the methods were originally developed to describe the current state of the weather in detail and to predict the expected change on a time scale of a few hours (cf. \cite{browning1989nowcasting,schmid2019nowcasting}). Nowcasting focuses on predictions for the present, the immediate future and the recent past, and is now being used in other fields, such as the economy, where important statistics on the current economic situation are available only after a considerable time lag (cf. \cite{banbura2013now}), and in medicine, where epidemic nowcasting is used during an ongoing epidemic outbreak to assess the current situation, taking into account the main pathogenic, epidemiological, clinical and socio-behavioural factors (cf. \cite{wu2021nowcasting}). Applications of nowcasting methods for econometrics can be found here in \cite{giannone2006nowcasting,fornaro2020nowcasting,bok2018macroeconomic} or for predictions of epidemic development here in \cite{johansson2014nowcasting,gunther2021nowcasting,birrell2021real}. When predicting values in the present, it is usual that the true values of the recent past are not available, so that only an estimate of these values is provided by the nowcast. Using the notation for forecasting above, this means that a prediction value $x_{t,\tau}$ also exiertits for $\tau$ less than or equal $t$. Similar to forecasting, nowcasters are often evaluated and  compared in the literature on the basis of performance measures such as the root mean absolute error (cf. \cite{gunther2021nowcasting}).

All the measures discussed, regardless of the area, have in common that the comparison and evaluation of the methods takes into account the local absolute differences in different ways, but does not directly consider the influence of the sign of the differences. To our best knowledge, there is no work in the literature that addresses the trending capability of measurement series or prediction with continuous outcomes. Because of the different ways in which data are generated and processed depending on the area of application, the data collected should be prepared in different ways for further processing in order to identify the trending ability (see chapter \ref{cap: trending}). In particular, the following shows how the time series of change (with lag $l$) should be defined, depending on whether measurement, forecasting or nowcasting methods are considered.

For measurement:
\begin{equation}
	\diffxl = (x_{t+l} -x_{t})^T_{t=0} \label{eq:diffxl_measure}
\end{equation}
For forecasting: 
\begin{equation}
	\diffxl = (x_{t,t+l} -y_{t})^T_{t=0} \label{eq:diffxl_forecasting}
\end{equation}
For nowcasting:
\begin{equation}
\diffxl = 
\begin{cases} 
(x_{t,t+l} -x_{t,t})^T_{t=0} & \text{if } y_{t} \text{ is not known at time } t, \\
(x_{t,t+l} -y_{t})^T_{t=0}  & \text{else}.
\end{cases} \label{eq:diffxl_nowcasting}
\end{equation}

\todo[inline]{Formeln als Tabelle oder in den Text einbinden.}
